---
title: "p8105_hw3_xm2229"
author: "Xiaoyue Ma"
date: "10/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

load the data set from the library
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

 - How many aisles are there, and which aisles are the most items ordered from?

```{r}
skimr::skim(instacart)
```

__The dataset is derived from the original "instacart" data. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. The structure of the data is `r class(instacart)`. Variables include `r names(instacart)`. Key variables descriptions can be find on__ http://p8105.com/dataset_instacart.html:

* order_id: order identifier
* product_id: product identifier
* add_to_cart_order: order in which each product was added to cart
* reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
* user_id: customer identifier
* eval_set: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train”  eval_set)
* order_number: the order sequence number for this user (1=first, n=nth)
* order_dow: the day of the week on which the order was placed
* order_hour_of_day: the hour of the day on which the order was placed
* days_since_prior_order: days since the last order, capped at 30, NA if order_number=1
* product_name: name of the product
* aisle_id: aisle identifier
* department_id: department identifier
* aisle: the name of the aisle
* department: the name of the department


__Here are some exmaples of some observations in from the dataset:__
```{r}
head(instacart, 5)
```

 - Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

__There are `r length(unique(instacart$aisle_id))` aisles and aisle #`r which.max(table(instacart$aisle_id))`, which is the `r unique(instacart[which(instacart$aisle_id == which.max(table(instacart$aisle_id))), ]$aisle)` aisle.__

```{r}
library(ggplot2)
instacart %>%
  count(aisle, name = "n_orders") %>%
  filter(n_orders >= 10000) %>%
  ggplot(aes(x = aisle, y= n_orders)) + 
  geom_bar(position = "dodge", stat = "identity") + coord_flip()
```

 - Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.


```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count), .by_group = TRUE) %>%
  group_by(aisle) %>%
  slice(1:3) %>%
  knitr::kable()
```

 - Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour) %>% # produce a 2*7 table
  knitr::kable(digits = 2)
```

# Problem 2

Load in the dataset
```{r}
data("brfss_smart2010")
```

First, do some data cleaning:

 - format the data to use appropriate variable names;
 - focus on the “Overall Health” topic
 - include only responses from “Excellent” to “Poor”
 - organize responses as a factor taking levels ordered from “Poor” to “Excellent”
 
```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))

head(brfss)
```
 
Using this dataset, do or answer the following (commenting on the results of each):

 - In 2002, which states were observed at 7 or more locations? What about in 2010?
 
```{r}
answer = brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarise(n = n_distinct(locationdesc)) %>% 
  filter(n >= 7)
```

__In 2002 `r answer$locationabbr` were observed at 7 or more locations.__

```{r}
answer_2010 = brfss %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  summarise(n = n_distinct(locationdesc)) %>% 
  filter(n >= 7)
```
 
__In 2010 `r answer_2010$locationabbr` were observed at 7 or more locations.__

 - Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


```{r}
brfss %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>%
  summarise(average_v = mean(data_value)) %>%
  ggplot(aes(x = locationabbr, y = average_v, color = year)) + 
  geom_line() +
  labs(
    title = "Mean data value in each state from 2002 to 2010",
    x = "State",
    y = "Number of locations"
  ) + coord_flip()
```

 - Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
 
```{r}
brfss %>%
  filter(locationabbr == "NY", year %in% c(2006, 2010)) %>%
  ggplot(aes(x = data_value, fill = response)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") +
  labs(
    title = "Density distribution of data_value for year 2006 and 2010",
    x = "data_value",
    y = "density"
  )+
  facet_grid(~year)

```

# Problem 3

load in the dataset
```{r}
accel <- read.csv('./data/accel_data.csv')
```

- Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_cleaned = accel %>% 
  janitor::clean_names() %>%
  mutate(weekdays = ifelse(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "weekday", "weekend")) %>%
  mutate(weekdays = forcats::fct_relevel(weekdays, c("weekday", "weekend")))
```

__The dataset is derived from the heart activity collected from an accelerometer of a patient in weeks. The dataset has `r nrow(accel_cleaned)` rows and `r ncol(accel_cleaned)` columns. The structure of the data is `r class(accel_cleaned)`. Key variables are: __

* week: number of week since the data collection begins
* day_id: day identifier
* day: day of the week when the data is collected
* weekdays: 1 if it is a weekday, 0 otherwise.
* activity_* : are the activity counts for each minute of a 24-hour day starting at midnight


- Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_cleaned %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "activity_sum"
  ) %>%
  group_by(day_id) %>%
  summarise(
    day_activity = sum(activity_sum)
    ) %>%
  knitr::kable()
```

__There is no apparent trend__

- Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r}
accel_cleaned %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "activity_sum"
  ) %>%
  group_by(day_id, day) %>%
  summarise(
    day_activity = sum(activity_sum)
    ) %>%
  ggplot(aes(x = day_id, y=day_activity, color=day)) + 
  geom_point() + 
  geom_line() + 
  labs(
    title = "24-hour activity for each day",
    x = "day",
    y = "24-hour activity"
  )
```

__According to the plot, it seems to be a peak activity in the first three weeks and a drop in the fourth and fifth week following a rise in the end.__